`You Draw It': Implementation of visually fitted trends with \texttt{r2d3}

### 2022-11-12 revisions/response-to-reviewers.qmd

<!-- \todo{Need to address yet.} \dubcheck{Double check / review.} -->

### Review: Associate Editor

This is a nicely written paper about an implementation of a tool to assess visually fitted trends. This tool was inspired by a New York Times feature "You Draw It" and was developed using the r2d3 R package. The authors explain the motivation for developing the tool, provide computing details, and present results from an experiment studying how people fit trends by visualizing them on a scatter plot.

The primary element missing from the paper is how this tool and insights from the experiment might be useful for the reader. One context that came to mind was classroom use, where instructors could use the tool in lessons about multivariable relationships and regression. It would be beneficial for the authors to include more detail about how this might be used in practice.

+ \response{We added a discussion section to address the appllicaiton of teaching and training by connecting the idea of the tool being used for interactive reading (NYT), interactive testing (our study), and interactive teaching/training (future applicaiton).}

@@

+ pg. 11 - Please provide a brief description of Prolific for readers unfamiliar with company.
  + \response{A breif description of prolific, "a crowd sourcing platform for researchers to recruit participants for their online studies," was added to the Validation Study section.}

### Review: Guest Editor for the SDSS Special Issue of the JDS

I agree with the AE’s decision of “Requires Major Revision”.

+ In addition to the AE’s comments, another possible application is using the tool for staff training. There is a growing demand to train up staff given the current job market.
  + \response{We added a discussion section to address the appllicaiton of teaching and training by connecting the idea of the tool being used for interactive reading (NYT), interactive testing (our study), and interactive teaching/training (future applicaiton).}

@@

+ Another comment is that I recommend the authors update figure 2 too (AE commented about figure 1), because the gray text is hard to see. 
  + \response{Figure 2 is a screenshot directly from the New York Times website used to demonstrate the use of the 'You Draw It' feature for interactive reading. While we could try to recreate this image or modify the colors, this changes the actual figure shown to readers in the New York Times.}

+ Also, figures 3 and 6 are difficult to read given there are three subplots. If there is space, perhaps the authors can reconfigure the plots, such as creating two rows (i.e., one plot in one row, and two plots in one row). 
  + \response{Figure 3 was reconstructed to provide the three subplots in vertical order spanning an entire page for better visability. As these subplots have a natural flow, we did not want to interupt this by creating a 2 + 1 subplot figure structure.}
  + \response{Figure 6 was reconstructed to provide a 1 + 2 subplot figure structure, keeping the two predicted plots from different models next to each other in the second row.}

+ Further, please note that all figures should be generated as vector images to ensure they’re high quality for final journal publication.
  + \response{Figure 1, Figure 5, and Figure 6 were generated within R and saved directly to pdf files from within the R environment. We expect these are all vector compatible; please let us know if this isn't sufficient or if we should use a different graphics device.}
  + \response{Figure 2 and Figure 3 were screenshots taken from websites. We zoomed in on a high quality screen and took a screenshot; saved as high resolution png files.}
  + \response{Figure 4 was sketched using an online tool; saving this as an svg changes the appearance of the font type and we would like to keep the sketched font. We saved this at the highest resoultion the tool allowed as a png file.}

### 2022-11-12 2022-sdss-you-draw-it-manuscript.qmd
added discussion section. needs to be reviewed/edited.

eyefitting_example <- eyefitting_model_data %>%
  # filter(participant_id == "65c10d498eae365e108efcd3dcb75287", parm_id == "N") %>%
  filter(participant_id == "60b16b9bd5a122c1457d31055df51a45", `Parameter Choice` == "F") %>%
  ggplot(aes(x = x)) +
  geom_line(aes(y = yols, color = "OLS", linetype = "OLS")) +
  geom_line(aes(y = ypca, color = "PCA", linetype = "PCA")) +
  geom_line(aes(y = ydrawn, color = "Drawn", linetype = "Drawn")) +
  geom_point(data = eyefitting_simulated_data %>%
               filter(dataset == "point_data", participant_id == "60b16b9bd5a122c1457d31055df51a45", `Parameter Choice` == "F"),
             aes(x = x, y = y)) +
  facet_wrap(~`Parameter Choice`, labeller = labeller(`Parameter Choice` = label_both)) +
  theme_bw(base_size = 14) +
  theme(aspect.ratio = 1,
        legend.position = "bottom",
        plot.title   = element_text(size = 14, hjust = 0),
        axis.text    = element_text(size = 12),
        axis.title   = element_text(size = 12),
        legend.title = element_text(size = 12),
        legend.text  = element_text(size = 12),
        # strip.text = element_text(size = 8, margin = margin(0.1,0,0.1,0, "cm")),
        # strip.background = element_rect(size = 0.8),
        legend.key.size = unit(1, "line")
        ) +
  scale_x_continuous(limits = c(0,20)) +
  scale_color_manual("", values = c("black", "steelblue", "orange")) +
  scale_linetype_manual("", values = c("dashed", "solid", "solid")) +
  scale_y_continuous("y")

@@

```{r gas-prices, echo = F, message = F, warning = F}
#| fig-cap: "Weekly average gas prices in the United States, 2019-June 2022. Additional mental operations a viewer might perform while looking at the plot are annotated in grey [@usenergyinformationadministrationWeeklyAllGrades2022]."
#| out-width: \columnwidth
#| label: fig-gas-prices
#| fig-res: 300
#| fig-width: 8
#| fig-height: 4
#| fig-env: figure

@@

annotations <- tibble(
  xmin = as.POSIXct(ymd(c("2020-03-15", "2022-02-26"))),
  xmax = as.POSIXct(ymd(c("2020-05-30", "2022-06-29"))),
  ymin = -Inf,
  ymax = Inf,
  label = c("COVID lockdowns", "War in Ukraine")
)
ggplot(gas, aes(x = Date, y = Price)) + 
  geom_rect(data = annotations, aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax), alpha= 0.2, fill = "grey", inherit.aes = F) + 
  geom_text(data = annotations, aes(x = xmin, y = 1, label = label), color = "grey") + 
  ylab("Price per Gallon ($)") + 
  ggtitle("Weekly Average Gas Prices in the United States, 2019 - 2022", subtitle = "US Energy Information Administration") + 
  geom_point() + 
  scale_x_datetime(limits = as.POSIXct(ymd("2019-01-01", "2022-06-29"))) + 
  geom_line(se = F, color = "grey", method = "loess", span = .1, alpha = .5, stat = "smooth", size = 1.5)
```

@@

![You Draw It task plot as shown to user.    
**left:** Initial state, with instructions *Use your mouse to fill in the trend in the yellow box region.*    
**middle:** User view during task completion.    
**right:** Finished state](images/ydi-stimuli.png){#fig-you-draw-it-task-plot}

@@

![Sketch of underlying code for 'You Draw It', illustrating the data simulation conducted in R, the initial setup of the visual stimuli with D3 source code, along with the iterative process between the user interaction and plotting in Shiny. Once the user is done drawing the line, we saved the results of the drawn line to a database for analysis.](images/code-sketch-2.png){#fig-you-draw-it-code-sketch}

@@

Participants were recruited via Prolific in March 2022; while this study does utilize a convenience sample, as this is primarily a perceptual task, previous results have found few differences between expert and non-expert participants in this context [@vanderplas2015spatial].
The study was conducted and distributed via a Shiny application and participants completed the tasks on their own computer, with a computer mouse, and in an environment of their choosing.
The data from this study were collected to validate this method of graphical testing, with the hopes of providing a new tool to assess graphical perception interactively.

@@

```{r eyefitting-plots, warning = F, message = F}
#| fig-cap: "Visual analysis results for parameter choice set F. \\linebreak \\textbf{A.} Example of validation feedback data where three trend lines show the the OLS fitted, PCA fitted, and participant drawn values overlaid on the simulated data points. \\linebreak \\textbf{B \\& C.} Estimated trends of residuals (vertical deviation of participant drawn points from both the OLS (blue) and PCA (orange) fitted points) as fit by the (B) linear mixed model and (C) generalized additive mixed model. A random sample of 75 participants was selected to display the individual participant residuals behind the overall trend."
#| out-width: \columnwidth
#| label: fig-eyefitting-plots
#| fig-res: 300
#| fig-width: 12
#| fig-height: 6
#| fig-env: figure

(eyefitting_example + ggtitle("A. Example")) + 
  (eyefitting_lmer_plot + ggtitle("B. LMER Results")) + 
  (eyefitting_gamm_plot + ggtitle("C. GAMM Results")) +
  plot_layout(ncol =  3)
```

@@

>>>>

eyefitting_example <- eyefitting_model_data %>%
  # filter(participant_id == "65c10d498eae365e108efcd3dcb75287", parm_id == "N") %>%
  filter(participant_id == "60b16b9bd5a122c1457d31055df51a45", `Parameter Choice` == "F") %>%
  ggplot(aes(x = x)) +
  geom_line(aes(y = yols, color = "OLS", linetype = "OLS")) +
  geom_line(aes(y = ypca, color = "PCA", linetype = "PCA")) +
  geom_line(aes(y = ydrawn, color = "Drawn", linetype = "Drawn")) +
  geom_point(data = eyefitting_simulated_data %>%
               filter(dataset == "point_data", participant_id == "60b16b9bd5a122c1457d31055df51a45", `Parameter Choice` == "F"),
             aes(x = x, y = y)) +
  facet_wrap(~`Parameter Choice`, labeller = labeller(`Parameter Choice` = label_both)) +
  theme_bw(base_size = 14) +
  theme(aspect.ratio = 1,
        legend.position = "right",
        plot.title   = element_text(size = 14, hjust = 0),
        axis.text    = element_text(size = 12),
        axis.title   = element_text(size = 12),
        legend.title = element_text(size = 12),
        legend.text  = element_text(size = 12),
        # strip.text = element_text(size = 8, margin = margin(0.1,0,0.1,0, "cm")),
        # strip.background = element_rect(size = 0.8),
        legend.key.size = unit(1, "line")
        ) +
  scale_x_continuous(limits = c(0,20)) +
  scale_color_manual("", values = c("black", "steelblue", "orange")) +
  scale_linetype_manual("", values = c("dashed", "solid", "solid")) +
  scale_y_continuous("y")

@@

```{r gas-prices, echo = F, message = F, warning = F}
#| fig-cap: "Demonstration of a viewer's engagement with a scatterplot showing the weekly average gas prices in the United States, from 2019-June 2022. The veiwer might perform several cognitive operations such as assessing points to determine outliers, fitting a rough mental smooth/trend line (overlaying solid gray line), and using additional contextual information from long term memory to explain the variation (shaded rectangular gray regions annotated in black text) [@usenergyinformationadministrationWeeklyAllGrades2022]. "
#| out-width: \columnwidth
#| label: fig-gas-prices
#| fig-res: 300
#| fig-width: 8
#| fig-height: 4
#| fig-env: figure

@@

annotations <- tibble(
  xmin = as.POSIXct(ymd(c("2020-03-15", "2022-02-26"))),
  xmax = as.POSIXct(ymd(c("2020-05-30", "2022-06-29"))),
  ymin = -Inf,
  ymax = Inf,
  label = c("COVID lockdowns", "War in Ukraine")
)
ggplot(gas, aes(x = Date, y = Price)) + 
  geom_rect(data = annotations, aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax), alpha= 0.3, fill = "gray30", inherit.aes = F) + 
  geom_text(data = annotations, aes(x = xmin, y = 1, label = label), color = "black") + 
  ylab("Price per Gallon ($)") + 
  ggtitle("Weekly Average Gas Prices in the United States, 2019 - 2022", subtitle = "US Energy Information Administration") + 
  geom_point(shape = 1) + 
  scale_x_datetime(limits = as.POSIXct(ymd("2019-01-01", "2022-06-29"))) + 
  geom_line(se = F, color = "grey30", method = "loess", span = .1, alpha = .6, stat = "smooth", size = 1.5)
```

@@

<!-- ![You Draw It task plot as shown to user.     -->
<!-- **left:** Initial state, with instructions *Use your mouse to fill in the trend in the yellow box region.*     -->
<!-- **middle:** User view during task completion.     -->
<!-- **right:** Finished state](images/ydi-stimuli.png){#fig-you-draw-it-task-plot} -->

```{r ydi-stimuli}
#| fig-cap: "You Draw It task plot as shown to user. \\textbf{top:} Initial state, with instructions, \\textit{Use your mouse to fill in the trend in the yellow box region.} \\textbf{middle:} User view during task completion. \\textbf{bottom:} Finished state."
#| out-width: "70%"
#| label: fig-you-draw-it-task-plot
#| fig-res: 300
#| fig-width: 3.3
#| fig-height: 10
#| fig-env: figure
#| warning: false
#| message: false

library(cowplot)
frame1 <- ggdraw() + draw_image("images/ydi-stimuli-frame1.PNG")
frame2 <- ggdraw() + draw_image("images/ydi-stimuli-frame2.PNG")
frame3 <- ggdraw() + draw_image("images/ydi-stimuli-frame3.PNG")

frame1 + frame2 + frame3 + plot_layout(ncol = 1)
```

@@

```{r, code sketch}
#| fig-cap: "Sketch of underlying code for 'You Draw It', illustrating the data simulation conducted in R, the initial setup of the visual stimuli with D3 source code, along with the iterative process between the user interaction and plotting in Shiny. Once the user is done drawing the line, we saved the results of the drawn line to a database for analysis."
#| out-width: \columnwidth
#| label: fig-you-draw-it-code-sketch
#| fig-res: 300
#| fig-width: 12
#| fig-height: 6
#| fig-env: figure
knitr::include_graphics("images/code-sketch.png")
# knitr::include_graphics("images/code-sketch.svg")
```

@@

Participants were recruited via Prolific, a crowd sourcing platform for researchers to recruit participants for their online studies, in March 2022; while this study does utilize a convenience sample, as this is primarily a perceptual task, previous results have found few differences between expert and non-expert participants in this context [@vanderplas2015spatial].
The study was conducted and distributed via a Shiny application and participants completed the tasks on their own computer, with a computer mouse, and in an environment of their choosing.
The data from this study were collected to validate this method of graphical testing, with the hopes of providing a new tool to assess graphical perception interactively.

@@

```{r eyefitting-plots, warning = F, message = F}
#| fig-cap: "Visual analysis results for parameter choice set F. \\linebreak \\textbf{A.} Example of validation feedback data where three trend lines show the the OLS fitted, PCA fitted, and participant drawn values overlaid on the simulated data points. \\linebreak \\textbf{B \\& C.} Estimated trends of residuals (vertical deviation of participant drawn points from both the OLS (blue) and PCA (orange) fitted points) as fit by the (B) linear mixed model and (C) generalized additive mixed model. A random sample of 75 participants was selected to display the individual participant residuals behind the overall trend."
#| out-width: \columnwidth
#| label: fig-eyefitting-plots
#| fig-res: 300
#| fig-width: 9
#| fig-height: 11
#| fig-env: figure

((eyefitting_example + 
    ggtitle("A. Example") + 
    theme(plot.title = element_text(hjust = 1/3), legend.position = "bottom"))
 ) /
 (((eyefitting_lmer_plot + 
      ggtitle("B. LMER Results")  +
     theme(legend.position = "bottom")
    ) + 
  (eyefitting_gamm_plot + 
     ggtitle("C. GAMM Results")  +
     theme(legend.position = "bottom")
   )
  )
  # plot_layout(guides = 'collect')
  ) +
  plot_layout(widths = c(1, 2))
```

@@

# Discussion

The 'You Draw It' graphical task presented in this work is an example of an interactive graphic. 
While any communication between user and device can be considered an interaction, interactive graphics should be defined as a user-driven direct manipulation of plots and plotting elements with an immediate reaction (Unwin 1999, Becker et al. 1987, Eick & Wills 1995).
Here we discuss the use of 'You Draw It' as a tool for interactive reading, interactive testing, and interactive teaching as a way to increase user experience and user engagement with data.
The New York Times first introduced the 'You Draw It' feature as a tool for interactive reading.
The inclusion of interactive graphics in news and media have been used to inform readers in efforts to reduce misconceptions and aid in storytelling [@geidner2015role; @george2011interactive; @george2013storytelling].
In our work, we demonstrated the use of the 'You Draw It' method as a tool for interactive testing of graphics.
We collect data of the user drawn lines and compare to statistical results in order to understand perceptual biases.
While this work has many possible directions for future work, there is a strong application to extend the 'You Draw It' tool for interactive teaching.
The line of best fit is introduced in secondary schools as a student's first encounter with the concept of statistical association [@casey2016students].
@casey2015examining and @casey2016finding conducted task based interviews to examine student's conceptions of the line of best fit prior to introducing the topic.
The implications of using interactive teaching tools for the line of best fit can help students make connections between the shift in slope and the location of the best fit line [@nagle2017slope].
@casey2015teachers conducted similar task based interviews wish pre-service mathematics teachers in order to understand the implications regarding teacher preparation for the teaching of this topic. 
This demonstrates the use of the 'You Draw It' interactive tool for training purposes both within a classroom setting and for those in the workforce.

### 2022-11-12 2022-sdss-you-draw-it-manuscript.tex
added discussion section. needs to be reviewed/edited.

\author[1]{Emily A. Robinson\thanks{Corresponding author. Email: erobin17@calpoly.edu}}
\author[2]{Reka Howard\footnote{Email: rekahoward@unl.edu}}
\author[2]{Susan VanderPlas\footnote{Email: susan.vanderplas@unl.edu}}
\affil[1]{Department of Statistics, CalPoly-San Luis Obispo}
\affil[2]{Department of Statistics, University of Nebraska-Lincoln}
\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[many]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={`You Draw It': Implementation of visually fitted trends with r2d3},
  pdfkeywords={graphics, user interaction, regression},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{`You Draw It': Implementation of visually fitted trends with
\texttt{r2d3}}
% \author{}
\date{}

\begin{document}
\maketitle
\begin{abstract}
How do statistical regression results compare to intuitive, visually
fitted results? Fitting lines by eye through a set of points has been
explored since the 20th century. Common methods of fitting trends by eye
involve maneuvering a string, black thread, or ruler until the fit is
suitable, then drawing the line through the set of points. In 2015, the
New York Times introduced an interactive feature, called `You Draw It,'
where readers are asked to input their own assumptions about various
metrics and compare how these assumptions relate to reality. This
research is intended to implement `You Draw It', adapted from the New
York Times, as a way to measure the patterns we see in data. In this
paper, we describe the adaptation of an old tool for graphical testing
and evaluation, eye-fitting, for use in modern web-applications suitable
for testing statistical graphics. We present an empirical evaluation of
this testing method for linear regression, and briefly discuss an
extension of this method to non-linear applications.
\end{abstract}
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, interior hidden, breakable, enhanced, sharp corners, borderline west={3pt}{0pt}{shadecolor}, boxrule=0pt]}{\end{tcolorbox}}\fi

@@

\caption{\label{fig-gas-prices}Weekly average gas prices in the United
States, 2019-June 2022. Additional mental operations a viewer might
perform while looking at the plot are annotated in grey (US Energy
Information Administration 2022).}

@@

{\centering \includegraphics{images/ydi-stimuli.png}

@@

\caption{\label{fig-you-draw-it-task-plot}You Draw It task plot as shown
to user.\\
\textbf{left:} Initial state, with instructions \emph{Use your mouse to
fill in the trend in the yellow box region.}\\
\textbf{middle:} User view during task completion.\\
\textbf{right:} Finished state}

@@

{\centering \includegraphics{images/code-sketch-2.png}

@@

Participants were recruited via Prolific in March 2022; while this study
does utilize a convenience sample, as this is primarily a perceptual
task, previous results have found few differences between expert and
non-expert participants in this context (VanderPlas and Hofmann 2015).
The study was conducted and distributed via a Shiny application and
participants completed the tasks on their own computer, with a computer
mouse, and in an environment of their choosing. The data from this study
were collected to validate this method of graphical testing, with the
hopes of providing a new tool to assess graphical perception
interactively.

@@

@@

@@

@@

>>>>

\usepackage[dvipsnames]{xcolor} % colors
\newcommand{\ear}[1]{{\textcolor{blue}{#1}}}
\newcommand{\svp}[1]{{\textcolor{RedOrange}{#1}}}
\newcommand{\rh}[1]{{\textcolor{Green}{#1}}}
\author[1]{Emily A. Robinson\thanks{Corresponding author. Email: erobin17@calpoly.edu}}
\author[2]{Reka Howard\footnote{Email: rekahoward@unl.edu}}
\author[2]{Susan VanderPlas\footnote{Email: susan.vanderplas@unl.edu}}
\affil[1]{Department of Statistics, CalPoly-San Luis Obispo}
\affil[2]{Department of Statistics, University of Nebraska-Lincoln}
\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[many]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={`You Draw It': Implementation of visually fitted trends with r2d3},
  pdfkeywords={graphics, user interaction, regression},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{`You Draw It': Implementation of visually fitted trends with
\texttt{r2d3}}
\author{}
\date{}

\begin{document}
\maketitle
\begin{abstract}
How do statistical regression results compare to intuitive, visually
fitted results? Fitting lines by eye through a set of points has been
explored since the 20th century. Common methods of fitting trends by eye
involve maneuvering a string, black thread, or ruler until the fit is
suitable, then drawing the line through the set of points. In 2015, the
New York Times introduced an interactive feature, called `You Draw It,'
where readers are asked to input their own assumptions about various
metrics and compare how these assumptions relate to reality. This
research is intended to implement `You Draw It', adapted from the New
York Times, as a way to measure the patterns we see in data. In this
paper, we describe the adaptation of an old tool for graphical testing
and evaluation, eye-fitting, for use in modern web-applications suitable
for testing statistical graphics. We present an empirical evaluation of
this testing method for linear regression, and briefly discuss an
extension of this method to non-linear applications.
\end{abstract}
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[breakable, enhanced, interior hidden, sharp corners, borderline west={3pt}{0pt}{shadecolor}, frame hidden, boxrule=0pt]}{\end{tcolorbox}}\fi

@@

\caption{\label{fig-gas-prices}Demonstration of a viewer's engagement
with a scatterplot showing the weekly average gas prices in the United
States, from 2019-June 2022. The veiwer might perform several cognitive
operations such as assessing points to determine outliers, fitting a
rough mental smooth/trend line (overlaying solid gray line), and using
additional contextual information from long term memory to explain the
variation (shaded rectangular gray regions annotated in black text) (US
Energy Information Administration 2022).}

@@

{\centering \includegraphics[width=0.7\textwidth,height=\textheight]{./images/fig-you-draw-it-task-plot-1.pdf}

@@

\caption{\label{fig-you-draw-it-task-plot}You Draw It task plot as shown
to user. \textbf{top:} Initial state, with instructions,
\textit{Use your mouse to fill in the trend in the yellow box region.}
\textbf{middle:} User view during task completion. \textbf{bottom:}
Finished state.}

@@

{\centering \includegraphics{images/code-sketch.png}

@@

Participants were recruited via Prolific, a crowd sourcing platform for
researchers to recruit participants for their online studies, in March
2022; while this study does utilize a convenience sample, as this is
primarily a perceptual task, previous results have found few differences
between expert and non-expert participants in this context (VanderPlas
and Hofmann 2015). The study was conducted and distributed via a Shiny
application and participants completed the tasks on their own computer,
with a computer mouse, and in an environment of their choosing. The data
from this study were collected to validate this method of graphical
testing, with the hopes of providing a new tool to assess graphical
perception interactively.

@@

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

The `You Draw It' graphical task presented in this work is an example of
an interactive graphic. While any communication between user and device
can be considered an interaction, interactive graphics should be defined
as a user-driven direct manipulation of plots and plotting elements with
an immediate reaction (Unwin 1999, Becker et al.~1987, Eick \& Wills
1995). Here we discuss the use of `You Draw It' as a tool for
interactive reading, interactive testing, and interactive teaching as a
way to increase user experience and user engagement with data. The New
York Times first introduced the `You Draw It' feature as a tool for
interactive reading. The inclusion of interactive graphics in news and
media have been used to inform readers in efforts to reduce
misconceptions and aid in storytelling (Geidner et al. 2015;
George-Palilonis and Spillman 2011, 2013). In our work, we demonstrated
the use of the `You Draw It' method as a tool for interactive testing of
graphics. We collect data of the user drawn lines and compare to
statistical results in order to understand perceptual biases. While this
work has many possible directions for future work, there is a strong
application to extend the `You Draw It' tool for interactive teaching.
The line of best fit is introduced in secondary schools as a student's
first encounter with the concept of statistical association (Casey and
Nagle 2016). Casey (2015) and Casey (2016) conducted task based
interviews to examine student's conceptions of the line of best fit
prior to introducing the topic. The implications of using interactive
teaching tools for the line of best fit can help students make
connections between the shift in slope and the location of the best fit
line (Nagle, Casey, and Moore-Russo 2017). Casey and Wasserman (2015)
conducted similar task based interviews wish pre-service mathematics
teachers in order to understand the implications regarding teacher
preparation for the teaching of this topic. This demonstrates the use of
the `You Draw It' interactive tool for training purposes both within a
classroom setting and for those in the workforce.

@@

\leavevmode\vadjust pre{\hypertarget{ref-casey2015examining}{}}%
Casey, Stephanie A. 2015. {``Examining Student Conceptions of
Covariation: A Focus on the Line of Best Fit.''} \emph{Journal of
Statistics Education} 23 (1).

\leavevmode\vadjust pre{\hypertarget{ref-casey2016finding}{}}%
---------. 2016. {``Finding What Fits.''} \emph{Mathematics Teaching in
the Middle School} 21 (8): 484--91.

\leavevmode\vadjust pre{\hypertarget{ref-casey2016students}{}}%
Casey, Stephanie A, and Courtney Nagle. 2016. {``Students' Use of Slope
Conceptualizations When Reasoning about the Line of Best Fit.''}
\emph{Educational Studies in Mathematics} 92 (2): 163--77.

\leavevmode\vadjust pre{\hypertarget{ref-casey2015teachers}{}}%
Casey, Stephanie A, and Nicholas H Wasserman. 2015.
{``TEACHERS'KNOWLEDGE ABOUT INFORMAL LINE OF BEST FIT.''}
\emph{Statistics Education Research Journal} 14 (1): 8--35.

@@

\leavevmode\vadjust pre{\hypertarget{ref-geidner2015role}{}}%
Geidner, Nick, Ivanka Pjesivac, Iveta Imre, Ioana Coman, and Dzmitry
Yuran. 2015. {``The Role of Interactive Graphics in Reducing
Misperceptions in the Electorate.''} \emph{Visual Communication
Quarterly} 22 (3): 133--45.

\leavevmode\vadjust pre{\hypertarget{ref-george2011interactive}{}}%
George-Palilonis, Jennifer, and Mary Spillman. 2011. {``Interactive
Graphics Development: A Framework for Studying Innovative Visual Story
Forms.''} \emph{Visual Communication Quarterly} 18 (3): 167--77.

\leavevmode\vadjust pre{\hypertarget{ref-george2013storytelling}{}}%
---------. 2013. {``Storytelling with Interactive Graphics: An Analysis
of Editors' Attitudes and Practices.''} \emph{Visual Communication
Quarterly} 20 (1): 20--27.

@@

\leavevmode\vadjust pre{\hypertarget{ref-nagle2017slope}{}}%
Nagle, Courtney, Stephanie Casey, and Deborah Moore-Russo. 2017.
{``Slope and Line of Best Fit: A Transfer of Knowledge Case Study.''}
\emph{School Science and Mathematics} 117 (1-2): 13--26.
